The Wolf in Sheep's Clothing\\ Subversive Adversarial Agent Identification

### 2016-08-04 article.tex
Adds Distributed State Estimation Methodology with some edits and removals. Removes need for glossary package. Adds new images.

\documentclass[conference]{IEEEtran}
\usepackage[pdftex]{graphicx}
\graphicspath{{images/}}
\usepackage[backend=biber, style=ieee]{biblatex}
\addbibresource{jprefwolf.bib}
\addbibresource{mwrefwolf.bib}
\usepackage{amsmath} % For better math in tables and equations
\usepackage[final]{microtype} % Makes things look better in the end
\usepackage{glossaries} % Just here so I don't need to remove \gls{} from text when moving from my thesis
\input{glossary.tex}

@@

\subsection{State Estimation}
State estimation is a fundamental problem for robotics. Without knowing its location, a robot cannot be very useful. One common technique for robot localization is to use a \gls{kf}. The \gls{kf} is one of the most studied and most heavily utilized Bayesian filter \cite[39-81]{ProbabilisticRobotics}. At its core, a \gls{kf} represents the state of a linear system as a multivariate normal distribution. By representing the state of the system as a normal distribution the filter can compactly represent the degree of belief of any possible state of the system.

The application of the \gls{kf} to the task of robot localization has been studied intently \cite{Localization2003, Mohsin2014}. In \cite{Mohsin2014}, the \gls{kf} is explained to be so popular for three main reasons. First, it is optimal under certain assumptions, second that it is recursive which is memory efficient, and finally that it relies only on having knowledge of noisy sensor data and not being able to measure certain state variables directly.

However, the \gls{kf} is best used for a linear system. Thus, plain \glspl{kf} are not generally used for localization. The two most popular options for real-world localization are the \gls{ekf} and \gls{ukf} which both extend the \gls{kf} to better handle non-linearity. The \gls{ukf} was Julier and Uhlmann, and presented in \cite{Julier1997}. The \gls{ukf} has two primary advantages over the \gls{ekf}. The first is that it is more accurate than the \gls{ekf} and this accuracy difference increases as the system becomes less linear. And second that it is computationally less complex than the \gls{ekf}.

A \gls{ukf} was chosen to be implemented for this paper, however many of the statements here regarding the filters can be applied to \glspl{ekf} as well. Our chosen software package also makes it trivially easy to switch between a \gls{ukf} and \gls{ukf}. For the remainder of this article, we will use the generic term \glspl{kf}, demonstrating the fact that these statements apply to both \glspl{ukf} and \glspl{ekf}.

\subsection{Secure State Estimation}
Secure State Estimation requires a focus on all three of these pieces.

\subsubsection{Integrity}
The accuracy of a \gls{kf} when the sensors are functioning property is not in question, so we must look at when the sensors are not functioning properly. This leads to two possible scenarios, sensor failure and false data injection.

In a sensor failure scenario one or more of the sensors will stop providing all feedback. Luckily, this is something that \glspl{kf} can handle without problem. The \gls{kf} is always using all available information to generate its state estimate, and if one sensor fails the only result is that the accuracy of the filter will decrease. Obviously, if 100\% of sensors fail, the filter will become non-functional, but as long as at least one sensor remains operational the filter will not have difficulties. It is also worth nothing that, due to the design of the filter, the \gls{kf} does not have issues with sensors going in and out of a failure state. When information is available it is used, and when it is not available the sensor is ignored.

Knowing that sensor failure is not a huge risk for the \gls{kf}, sensor interference is a much more complex problem. If an attacker is somehow able to inject false data into the filter the pose estimate can be skewed. And this is a very real problem. The number of ways that false data could be added to a filter are too numerous to count, ranging from physically blocking a lidar sensor to hijacking the network packets of a wirelessly communicating sensor.

\cite{Mo2010} and \cite{Yang2013} both demonstrate that \glspl{kf} are susceptible to false data injection attacks, and that even with failure detectors a clever adversary can craft their attack to bypass these detectors. \cite{Bezzo_2014} and \cite{Mo2014} both attack this problem by adding extra steps into their \gls{kf} algorithm. Both successfully show that the filter output can be shielded from the effects of the attack.

\subsubsection{Availability}
In a distributed system, \gls{DoS} attacks become a real threat. While there are many ways to execute such attacks and also many ways to defeat them (\cite{wood2002denial}, \cite{bellardo2003802}) such systems are outside the scope of this TurtleBot. Rather than showing that the system can defeat \gls{DoS} attacks, we will show that it can function sufficiently in the event that there is a network interruption and inter-robot communication is prevented.

@@

Other work into distributed systems still makes the assumption of a cooperative team. \cite{Sanderson1997} and \cite{Roumeliotis2002} both demonstrate a system where a \gls{kf} is implemented that estimates the state of multiple robots in a system. However these states are computed relative to each other, considering the network of robots to be essentially one large robot with different parts. Continuing the highway example, cars will be moving in and out of communication with each other as some enter the highway and others exit. Thus the group of robots will be constantly changing and evolving, a challenge not tackled by these works.

The previously mentioned works are all examples of a decentralized, rather than distributed \gls{kf} problem. \cite{Olfati-Saber2005} makes clear the distinction between these two, saying:
\begin{quote}"...[decentralized] methods require a complete network with all-to-all links. This solution is not scalable for large-scale sensor networks... Thus, decentralized Kalman filtering and distributed Kalman filtering are two separate problems. In the latter one, each node only is allowed to communicate with its nieghobrs on a graph $G$ that is connected but rather sparse."
\end{quote}
\cite{Olfati-Saber2005} presents a solution to the distributed problem using consensus filters and multiple \glspl{kf}. However this work still requires adaptation of the \glspl{kf} and extra computation, which is something we are trying to avoid.

@@

\subsection{Distributed State Estimation}
[INSERT DISTRIBUTED STATE ESTIMATION METHODOLOGY HERE]

>>>>

\documentclass[conference]{IEEEtran}
\usepackage[pdftex]{graphicx}
\graphicspath{{images/}}
\usepackage[backend=biber, style=ieee]{biblatex}
\addbibresource{jprefwolf.bib}
\addbibresource{mwrefwolf.bib}
\usepackage{amsmath} % For better math in tables and equations
\usepackage[final]{microtype} % Makes things look better in the end
\usepackage[hidelinks]{hyperref}

@@

\subsection{State Estimation}
State estimation is a fundamental problem for robotics. Without knowing its location, a robot cannot be very useful. One common technique for robot localization is to use a Kalman Filter (KF). The KF is one of the most studied and most heavily utilized Bayesian filter \cite[39-81]{ProbabilisticRobotics}. At its core, a KF represents the state of a linear system as a multivariate normal distribution. By representing the state of the system as a normal distribution the filter can compactly represent the degree of belief of any possible state of the system.

The application of the KF to the task of robot localization has been studied intently \cite{Localization2003, Mohsin2014}. In \cite{Mohsin2014}, it is explained to be so popular for three main reasons. First, it is optimal under certain reasonable and common assumptions, second that it is recursive which is memory efficient, and finally that it relies only on having knowledge of noisy sensor data and not being able to measure certain state variables directly.

For this paper, we utilize an Unscented Kalman Filter (UKF) for the task of sensor fusion and localization. The UKF provides better accuracy for non-linear systems than a KF or Extended Kalman Filter (EKF) and is less computationally complex \cite{Julier1997}.

REMOVED EXPANDED EKF/UKF DISCUSSION

\subsection{Secure State Estimation}
Secure State Estimation requires a focus on all three of the previously mentioned facets of security: Integrity, Availability and Confidentiality. We choose to focus on the first and third of these, Integrity and Confidentiality, for our secure estimation system.

\subsubsection{Integrity}
The accuracy of a KF when the sensors are functioning properly is not in question, so we must look at when the sensors are not functioning properly. We propose two primary failure modes for data integrity: sensor failure and false data injection.

In a sensor failure scenario one or more of the sensors will stop providing all feedback. Luckily, this is something that KFs can handle without problem. The KF is always using all available information to generate its state estimate, and if one sensor fails the only result is that the accuracy of the filter will decrease. Obviously, if 100\% of sensors fail, the filter will become non-functional, but as long as at least one sensor remains operational the filter will not have difficulties. It is also worth nothing that, due to the design of the filter, the KF does not have issues with sensors going in and out of a failure state. When information is available it is used, and when it is not available the sensor is ignored.

Knowing that sensor failure is not a huge risk for the KF, sensor interference is a much more complex problem. If an attacker is somehow able to inject false data into the filter the pose estimate can be skewed. And this is a very real problem. The number of ways that false data could be added to a filter are too numerous to count, ranging from physically blocking a lidar sensor to hijacking the network packets of a wirelessly communicating sensor.

\cite{Mo2010} and \cite{Yang2013} both demonstrate that KFs are susceptible to false data injection attacks, and that even with failure detectors a clever adversary can craft their attack to bypass these detectors. \cite{Bezzo_2014} and \cite{Mo2014} both attack this problem by adding extra steps into their KF algorithm. Both successfully show that the filter output can be shielded from the effects of the attack.

REMOVED BRIEF ASIDE ON AVAILABILITY

@@

Other work into distributed systems still makes the assumption of a cooperative team. \cite{Sanderson1997} and \cite{Roumeliotis2002} both demonstrate a system where a KF is implemented that estimates the state of multiple robots in a system. However these states are computed relative to each other, considering the network of robots to be essentially one large robot with different parts. Continuing the highway example, cars will be moving in and out of communication with each other as some enter the highway and others exit. Thus the group of robots will be constantly changing and evolving, a challenge not tackled by these works.

The previously mentioned works are all examples of a decentralized, rather than distributed KF problem. \cite{Olfati-Saber2005} makes clear the distinction between these two, saying:
\begin{quote}"...[decentralized] methods require a complete network with all-to-all links. This solution is not scalable for large-scale sensor networks... Thus, decentralized Kalman filtering and distributed Kalman filtering are two separate problems. In the latter one, each node only is allowed to communicate with its nieghobrs on a graph $G$ that is connected but rather sparse."
\end{quote}
\cite{Olfati-Saber2005} presents a solution to the distributed problem using consensus filters and multiple KFs. However this work still requires adaptation of the KFs and extra computation, which is something we try to avoid.

@@

\subsection{Distributed State Estimation}
KFs perform better with more inputs. This is the driving force behind increasing the accuracy of mobile robot localization. With more sensor inputs, noise can be removed more accurately. One way to increase the number of sensor inputs to a filter is to add more sensors to the robot. However this will rapidly drive up cost and complexity. Another solution is to have robots share their information about the surrounding world, thus artificially inflating the number of sensors a robot has.

The flow for this system is as follows. A robot at position $p_l$, where $l$ represent that this position is with respect to the local frame, detects another robot at position $g_l$. The robot at $p_l$ then calculates its' pose in with respect to the global frame, $p_g$ and then transforms $g_l \rightarrow g_g$. This position $g_g$ is then shared with the robot's in the surrounding environment, which integrate the position into their own localization estimate. This is summarized in \autoref{tab:external_pose}.

\begin{table}
\centering
\begin{tabular}{c}
KF output $\rightarrow p_l, p_g$ \\
Lidar scan $+ p_l \rightarrow g_l$ \\
$g_g \rightarrow$ KF input
\end{tabular}
\caption{General algorithm for computing external pose.}
\label{tab:external_pose}
\end{table}

The following sections how the position estimates $p_l$, $p_g$, $g_l$, and $g_g$ are obtained. All localization estimates will be referenced with an associated coordinate frame. These frames are described in ROS REP-105 \cite{REP_105}. They are the base\_link (aka base\_footprint) frame, which is rigidly fixed to the center of the robot, the odom (aka local) frame, which is continuous but subject to drift over time, and the map (aka global) frame, which is subject to discrete jumps but does not drift over time.

REMOVED LENGTHY COORDINATE FRAME DESCRIPTIONS FROM HERE

\paragraph{Filters}
There are two filters operating on the robot to create a whole localization estimate. The first is the \textit{Continuous Filter} (CF), and the second is the \textit{Discrete Filter} (DF). They each perform a separate but related localization task, and combined give a full localization estimate with respect to both the local and global frame of the robot.

\subsubsection{Continuous Filter} \label{con_filter_subsubsection}
The CF calculates the transformation from $odom \rightarrow base\_link$. It receives odometry information from the robot's wheel encoders and IMU. Using this information it publishes the $odom \rightarrow base\_link$ transformation. Another name for the CF is the Self Filter.

\subsubsection{Discrete Filter} \label{disc_filter_subsubsection}
The DF calculates the transformation from $map \rightarrow base\_link$. Because the ROS TF package that handles coordinate frame transformations enforces a tree structure where each node can only have one parents, the DF references the $odom \rightarrow base\_link$ transform published by the CF and publishes a transform from $map \rightarrow odom$. The inputs to the DF are the robot's wheel encoder odometry, IMU readings, a GPS, and external poses calculated by other robots. Another name for the DF is the Distributed Filter.

REMOVED EKF VS UKF JUSTIFICATION FROM HERE

REMOVED SPECIFIC I/O EXPLANATION FOR FILTERS

REMOVED NODE STRUCTURE AND DETAILED NODE DESCRIPTIONS FROM HERE

REMOVED COMMUNICATIONS LOGIC JUSTIFICATION (I.E. PUBLISH SUBSCRIBE VS ACTION SERVER) FROM HERE

REMOVED SIMULATION DETAILS

REMOVED NOISE MODEL EXPLANATION
